Sketch Recognition using Particle Swarm Optimizations

### 2009-04-14 trunk/paper2/Paper2.tex
after my edit today 15 midnight

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. There are no detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
 \begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.45]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.45]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}
\begin{figure}[]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}
Dominant points are characterized by lower speed values, higher curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

@@

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two particle swarm algorithms that divide the stroke to either lines or lines and curves. The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.7]{images/blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsubsection{Ellipse Detection} 
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms.

\textit{The first algorithm \textsl{AlgS1}}\\
The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$. The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed using equation %\ref{eq:fitnessSwarm1}
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution. Otherwise the system favors the lower number of vertices.

 \textit{The second algorithm \textsl{AlgS2}} \\
The algorithm has the same problem formulation but different fitness and error functions are used. An attempt is made to fit each segment into line or circular arc. The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all segments is the total error of the particle.  The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
 \begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation %\ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. The larger $k$ is, the more effect the number of segments will have. For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.
\end{description}

After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. Otherwise, one of these points are removed based on the error computed by the segmentation algorithm.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The following list describes each feature in details.
\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[No. Of segments] Number of segments in the symbol. 
		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [No. of curves] Number of curves or ellipses in the symbol. 
		\item [No. of lines] Number of lines in the symbol. 
		\item [No. of perpendicular lines] Number of perpendicular lines. 
		\item [No. of parallel lines] Number of parallel lines. 
		\item [No. of intersections] Number of intersection between lines and curves. 
		\item [Size Ratio] Ratio between width to height of the symbol.
		%\item[No. of holes] Number of holes in the symbols.

\end{description}
\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. 
\item[Statistical Features (FS3)]   
\begin{description}
	\item[Zernike moments]  Zernike moments of order \cite{HeloiseBeautification}. 
\end{description}
	\item[Composite Features (FS4)] Features that are composites of different geometrical and statistical symbol. 
	\begin{description}

	\item [Ink density] Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item[Convex Hull Area] Area of convex hull with respect to area of bounding box of symbol.
	\item[Convex Hull Perimeter] Perimeter of convex hull with respect to total length of symbol.
		\item [Mean Centroidal radius] The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item [Mean Time difference] The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{description}

\end{description}
%The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. 
After computing the features the symbol is introduced to the classifier. The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object) is used to handle multi-class.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.

\section{Experimental}
\label{sec:Experiments}
The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.  
\begin{figure}
\centering 
\label{fig:symbolSet}
\fbox{ \parbox{5cm}{% 
		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
		\caption{The Symbol Set} Symbols used in the dataset.
\end{figure}
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
A set of test were performed on the system to enhance both the swarm algorithms outputs. The effect of number of agents, the error thresholds and various other paramters was investigated. The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm). The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time. Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).

@@

\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/testsym.jpg}
	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:test2}
\end{figure*}  
The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This lead us to the next experiment to choose best set of feature to recognize symbols. 
\begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 \include{conf}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}
\subsection{Features Analysis}
\label{sec:featexp}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
 \begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/testFeat.jpg}
	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure*}

 We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous. It is only based on what the user intends to draw while sketching the symbol. Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.4]{images/testF1.jpg}
	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}

@@

\section{Conclusion and Future Work}
\label{ConclusionandFutureWork}
This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information which help improving the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results. The use of statistical moments and structural features improves the recognition rate. The system was tested on 13 different symbols and achieved an accuracy of 93.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system process one symbol at a time, an enchantment would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}

>>>>

\subsection{Preprocessing}
\label{Prepross}% Agar \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena.
 Speed and time difference information were widely used in sketch recognition systems \cite{earlyprocess}. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}. It was noted that time difference provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mention that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals as long as the pointing device is moving. Therefore, there are no new detected points when the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary (Figure \ref{fig:timediff}). Contrary to speed information, where the users draw with variant speeds, the speed information has a lot of noise in the data.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}.% Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}
 \begin{figure}
	\centering
			\subfigure[ Speed Graph]{	\includegraphics[scale=0.4]{images/vel3.jpg}}
			\hfill
			\subfigure[ Time Difference Graph] {\label{fig:timediff}\includegraphics[scale=0.4]{images/td3.jpg}}
	\caption{Speed and Time Difference Graphs}
	\label{fig:speed2Distance}
\end{figure}
\begin{figure}[]
	\centering
		\includegraphics[scale=0.5]{images/stroke3.jpg}
	\caption{Input Stroke} Example of an input stroke to the system. 
	\label{fig:orignalStroke}
\end{figure}
Dominant points are characterized by lower speed values, higher curvature and direction values. However, the dominant points from the curvature information may not be the same points as the dominant points extracted from the speed information. Hence, we compute time difference, direction, speed and curvature of each point along a stroke to ensure that no dominant point is overlooked in this phase. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$.

@@

\subsection{Segmentation}
\label{seg}
In the segmentation stage the input stroke is divided into a set of primitives. As shown in Figure \ref{fig:segblock} first an attempt is made to fit the stroke points into a curve or an ellipse using a minimum square error fitting algorithm \cite{chernov-2003}. The ellipse fitting step helps to prevent the system from over segmenting the stroke into multiple lines or curves if the input stroke is an ellipse. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed to two furthur segmentation algorithms that divide the stroke to either lines or lines and curves. The next section describes in details the ellipse detection algorithm and the two PSO algorithms used to divide the stroke. % The input to these algorithms are the stroke points along with the possible dominant points $P_{pd}$ computed during preprocessing. The output is a set of dominant points which are connected by either lines or curves.

%///AllBlockDiagram.jpg  you can change the block to the main block. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/blockSmall.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\end{figure}
\subsection{Ellipse fitting} 
%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation \cite{chernov-2003}.  
t
This process tries to fit the stroke points into an ellipse arc; it starts with computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the center of the ellipse. The axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box. The least square fitting algorithm \cite{chernov-2003} is used to minimize the fitting error of the ellipse Equation (\ref{eq:circleFit})

@@

	\item[Fitness function:] The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. 
	\end{description}
\subsubsection{\textit{Polygon approximation\textsl{AlgS1}}}
%\\
The approximation error is computed by the equation \ref{eq:ErrorSwarm1}. The graphical meaning of the error is shown in Figure\ref{fig:LabelsPSO}.
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation} where the arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i$ to point $x_{j}$ as in $x_i,x_{i+1} \cdots,x_j$. The line $\overline{x_i x_j} $ as the straight line connecting point $x_i$ to point $x_j$, $M$ is the number of dominant points in this solution as generated by the swarm algorithm. The error $e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$. The fitness is computed using equation %\ref{eq:fitnessSwarm1}
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $\varepsilon$ the fitness is given a -ve value to lower the fitness value of the solution. Otherwise the system favors the lower number of vertices.

 \subsubsection{\textit{Hybird Fitting \textsl{AlgS2}}}
The algorithm has the same problem formulation but different fitness and error functions are used. An attempt is made to fit each segment into a line or circular arc. The errors of both circle and line fit estimations are computed for each segment $S_i=\widehat{x_ix_j}$. The approximation with the lower error value is chosen as the final approximation of this segment $S_i$\cite{CruveDivisionSwarm}. The sum of the approximation errors of all segments is the total error of the particle.  The total error of the particle is computed by equation %\ref{eq:errorSwarm2}
 \begin{equation}
E=\sum\nolimits_{i = 0}^M e(D_i) 
\label{eq:errorSwarm2}
\end{equation}where $M$ is the number of segments in the solution as generated by the swarm algorithm, $D_i$ is the minimum approximation error of curve and line approximations $min(d_c,d_l)$ where $d_c$ is the circle approximation error and $d_l$ is the line approximation error as computed by \cite{CruveDivisionSwarm}.  The fitness is computed by the equation %\ref{eq:fitnessSwarm2} 
\begin{equation}
\max fitness(P_i ) = \frac{1}{{E \times M^k }}
\label{eq:fitnessSwarm2}
\end{equation} where $E$ is the error and $M$ is the number of segments and $k$ is a parameter tweaked to get minimum number of segments. The larger $k$ is, the more effect the number of segments will have. For our system, $k$ is selected to be 0.5\cite{CruveDivisionSwarm}.

After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The following list describes each feature in details.
%\begin{description}
\subsubsection{Structural and geometrical Features(FS1)}
  Features defines the structure of the geometrical symbol.  
%\begin{description}
%	 \item[No. Of segments] Number of segments in the symbol. 
%		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [No. of curves] Number of curves or ellipses in the symbol. 
%		\item [No. of lines] Number of lines in the symbol. 
%		\item [No. of perpendicular lines] Number of perpendicular lines. 
%		\item [No. of parallel lines] Number of parallel lines. 
%		\item [No. of intersections] Number of intersection between lines and curves. 
%		\item [Size Ratio] Ratio between width to height of the symbol.
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
 \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
\subsubsection{Rubine Feature Set (FS2)}
  Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. Feature such as length of stroke, smoothness of symbol and properties of the bounding box of the symbol.

\subsubsection{Statistical Features (FS3)}  
\begin{itemize}
	\item \emph{Zernike moments } Zernike moments of order \cite{HeloiseBeautification}. 
\end{itemize}
\subsubsection{Composite Features (FS4)}
 Features that are composites of different geometrical and statistical symbol. 
	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{itemize}

%\end{description}
%The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. 
After computing the features the symbol is introduced to the classifier. The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object) is used to handle multi-class.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.

\section{Experimental}
\label{sec:Experiments}
The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. The results displayed are the average recognition accuracy of the four splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.  
\begin{figure}
\centering 
\label{fig:symbolSet}
\fbox{ \parbox{5cm}{% 
		\includegraphics[scale=0.5]{images/symbolSet.PNG}	}}
		\caption{The Symbol Set} Symbols used in the dataset.
\end{figure}
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
A set of tests were performed on the system to enhance non ellipse fitting algorithm. The effect of number of agents, the error thresholds and various other paramters were investigated. The result of these tests are shown in Figure \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increases the error value decrease (error of the segmentation algorithm). The final number of agents and maximum swarm iteration used in the system was based on a trade off between error achieved versus the computation time. Other algorithm parameters such as $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}).

@@

\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/testsym.jpg}
	\caption{Symbols Comparison} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:test2}
\end{figure*}  
The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This lead us to the next experiment to choose best set of feature to recognize symbols. 
\begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 \include{conf}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}
\subsection{Features Analysis}
\label{sec:featexp}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
 \begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/testFeat.jpg}
	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure*}  
 We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous. It is only based on what the user intends to draw while sketching the symbol. 
 The main features affected by teh segmentation stage are the geometeric features computed in\textbf{FS1 }. Hence those are the features we used to estimate the efficency of the segmentation algorithm. The recognition accuracy using only the gometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.4]{images/testF1.jpg}
	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}

@@

\section{Conclusion and Future Work}
\label{ConclusionandFutureWork}
This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information which help improving the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results. The use of statistical moments and structural features improves the recognition rate. The system was tested on 13 different symbols and achieved an overall accuracy of 93.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Another area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}
